---
title: "Dataset Examples"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set( echo = TRUE, message=F, warning=F )

library( stargazer )
library( AER )
library( wooldridge )
library( Ecdat )
library( dplyr )
library( memisc )
library( pander )


jplot <- function( x1, x2, cex=3.5, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.2), 
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}



panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if( missing(cex.cor) ) cex <- 0.8/strwidth(txt)
    
    test <- cor.test(x,y)
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " "))
    
    text(0.5, 0.5, txt, cex = 1.5 )
    text(.7, .8, Signif, cex=cex, col=2)
}




```


```{r, eval=F, echo=F}


# Stargazer Example

stargazer( m.103, m.203, m.204, m.205, m.207, out="./Results/RegressionModel.doc",
           dep.var.labels.include=FALSE, dep.var.caption="DV: Grant Recipient is Local [1,0]",
           column.labels=c("Global","Asia","Africa","Latin America","Middle East"),
           omit.stat = c("rsq", "f", "ser"), 
           omit=c("GM_KEY","REGION_BY_SECTOR","GRANT_REGION_IMPUTED","PRIMARY_CODE_DESC"), 
           add.lines = list(c("DV: Proportion Local",prop.local),
                            c("Program Area Fixed Effects?","Yes","Yes","Yes","Yes","Yes","Yes")),
           covariate.labels=c("Foundation Age",
                              "Foundation Size (log of assets)", 
                              "Foundation Size Squared",
                              "Foundation Regional Experience",
                              "Foundation Program Area Experience",
                              #"Prior NGO Experience",
                              "Grant Size (log of amount)",
                              "Grant Length",
                              "Prop Grants to Foreign NGOs",
                              "Prop. of Grants Int.",
                              "Intercept"),
           digits=2, type="html" )

```







# Package *AES*

Applied Econometrics with R

Functions, data sets, examples, demos, and vignettes for the book Christian Kleiber and Achim Zeileis (2008),
Applied Econometrics with R, Springer-Verlag, New York.


```{r, eval=F}

install.packages( "AER" )
library( AER )

vignette( "AER" )

```

## California Schools Data

The data used here are from all 420 K-6 and K-8 districts in California with data available for 1998
and 1999. 

The dataset is used to address the question of whether 

Test scores are on the Stanford 9 standardized test administered to 5th grade students.
School characteristics (averaged across the district) include enrollment, number of teachers (measured
as “full-time equivalents”, number of computers per classroom, and expenditures per student.
Demographic variables for the students are averaged across the district. The demographic variables
include the percentage of students in the public assistance program CalWorks (formerly AFDC),
the percentage of students that qualify for a reduced price lunch, and the percentage of students that
are English learners (that is, students for whom English is a second language).


**Data Dictionary**
 
* **district**: District code (character) 
* **stratio**: Student-teacher ratio (calculated below) 
* **score**: Average of two primary scores, math and reading 
* **school**: School name (character) 
* **county**: indicating county (factor) 
* **grades**: indicating grade span of district (factor) 
* **students**: Total enrollment 
* **teachers**: Number of teachers 
* **calworks**: Percent qualifying for CalWorks (income assistance) 
* **lunch**: Percent qualifying for reduced-price lunch 
* **computer**: Number of computers 
* **expenditure**: Expenditure per student 
* **income**: District average income (in USD 1,000) 
* **english**: Percent of English learners 
* **read**: Average reading score 
* **math**: Average math score 
  



### Specification

What happens to the significance of the student-teacher ratio variable once appropriate controls are added to the model? 



```{r, echo=F}

panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y, use="pairwise.complete.obs")
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
    
    test <- cor.test(x,y)
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " "))
    
    text(0.5, 0.5, txt, cex = 2 )
    text(.7, .8, Signif, cex=3, col=2)
}


panel.smooth <- function (x, y, col = par("col"), bg = NA, pch = par("pch"), 
  cex = 1, col.smooth = "red", span = 2/3, iter = 3, ...) 
{
  points(x, y, pch = 19, col = gray(0.5,0.5), 
         bg = bg, cex = 0.7)
  ok <- is.finite(x) & is.finite(y)
  if (any(ok)) 
    lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
      col = col.smooth, lwd=2, ...)
}



```




```{r, fig.width=10, fig.height=10}


data( "CASchools" )

## variable transformations
CASchools$stratio <- with(CASchools, students/teachers)
CASchools$score <- with(CASchools, (math + read)/2)


dat <- dplyr::select( CASchools, score, stratio, expenditure, english, lunch, calworks )

pairs( dat, lower.panel=panel.smooth, upper.panel=panel.cor)

```


```{r, results='asis'}

## Stock and Watson (2007)
## p. 152
fm1 <- lm( score ~ stratio, data = CASchools )
# coeftest(fm1, vcov = sandwich)

## p. 159
fm2 <- lm( score ~ I(stratio < 20), data = CASchools )

## p. 199
fm3 <- lm( score ~ stratio + english, data = CASchools )

## p. 224
fm4 <- lm( score ~ stratio + expenditure + english, data = CASchools )


stargazer( fm1, fm2, fm3, fm4, type="html",
           omit.stat = c("rsq", "f", "ser"),
           digits=2)

```




### Additional Controls

Control for non-native speakers and income level of students. 

```{r}
## Table 7.1, p. 242 (numbers refer to columns)
fmc3 <- lm( score ~ stratio + english + lunch, data = CASchools )
fmc4 <- lm( score ~ stratio + english + calworks, data = CASchools )
fmc5 <- lm( score ~ stratio + english + lunch + calworks, data = CASchools )

```


```{r, results='asis'}

stargazer( fmc3, fmc4, fmc5, type="html",
           omit.stat = c("rsq", "f", "ser"),
           digits=2 )

```







## Teacher Ratings

```{r}

data("TeachingRatings")

TeachingRatings$response.rate <- TeachingRatings$students / TeachingRatings$allstudents

TeachingRatings %>% head() %>% pander()

```




```{r, fig.width=10, fig.height=10}

dat <- dplyr::select( TeachingRatings, eval, age, beauty, response.rate )

pairs( dat, lower.panel=panel.smooth, upper.panel=panel.cor)

```


* Are beautiful people better teachers? 
* What is the relationship between age and response rate? 
* Are people content with a course more or less likely to submit evaluations? 



```{r}

par( mfrow=c(2,2) )
plot( TeachingRatings$gender, TeachingRatings$eval, frame.plot=F, outline=F, main="Gender" )
plot( TeachingRatings$minority, TeachingRatings$eval, frame.plot=F, outline=F, main="Minority" )
plot( TeachingRatings$tenure, TeachingRatings$eval, frame.plot=F, outline=F, main="Tenure Status" )
plot( TeachingRatings$credits, TeachingRatings$eval, frame.plot=F, outline=F, main="Credits" )


```





```{r}

group.structure <- formula( TeachingRatings$eval ~ TeachingRatings$minority * TeachingRatings$gender )

boxplot( group.structure, ylim=c(3,5),
         las=0, frame.plot=F, outline=F, 
         main="Performance by Gender & Minority Status",
         col=c("steelblue4","steelblue3","firebrick4","firebrick3"), 
         staplewex=0, whisklty=0, border="gray90", lwd=2 )

abline( h=seq(3,5,0.25), col="gray", lty=3, lwd=0.25 )

```



```{r}

group.structure <- formula( TeachingRatings$eval ~ TeachingRatings$gender * TeachingRatings$tenure )

boxplot( group.structure, ylim=c(3,5),
         las=0, frame.plot=F, outline=F, 
         main="Performance by Gender and Tenure Status of the Professor",
         col=c("steelblue","firebrick","steelblue","firebrick" ), 
         staplewex=0, whisklty=0, border="gray90", lwd=2 )

abline( h=seq(3,5,0.25), col="gray", lty=3, lwd=0.25 )

```



```{r}

## evaluation score vs. beauty
jplot( TeachingRatings$beauty, TeachingRatings$eval, xlab="Beauty", ylab="Student Evaluations" )
fm <- lm(eval ~ beauty, data = TeachingRatings)
abline(fm)
summary(fm)
## prediction of Stock & Watson's evaluation score
sw <- with(TeachingRatings, mean(beauty) + c(0, 1) * sd(beauty))
names(sw) <- c("Watson", "Stock")
predict(fm, newdata = data.frame(beauty = sw))

## Hamermesh and Parker, 2005, Table 3

fmw <- lm(eval ~ beauty + gender + minority + native + tenure + division + credits,
weights = students, data = TeachingRatings)
coeftest(fmw, vcov = sandwich)
## (same coefficients but with different covariances)
```




```{r, results='asis'}

stargazer( fmc3, fmc4, fmc5, type="html",
           omit.stat = c("rsq", "f", "ser"),
           digits=2 )

```





# Package 'woolridge'

Read directly from the website:

```{r, eval=F}

library( foreign )

dat <- read.dta( "http://fmwww.bc.edu/ec-p/data/wooldridge/meap01.dta" )

dat <- read.dta( "http://fmwww.bc.edu/ec-p/data/wooldridge/affairs.dta" )



```

R package:

```{r, eval=F}

install.packages( "wooldridge" )
library( wooldridge )

```




## Affairs Dataset

```{r}

data( "affairs" )
head( affairs ) %>% pander()

```


* id: identifier 
* male: =1 if male 
* age: in years 
* yrsmarr: years married 
* kids: =1 if have kids 
* relig: 5 = very relig., 4 = somewhat, 3 = slightly, 2 = not at all, 1 = anti 
* educ: years schooling 
* occup: occupation, reverse Hollingshead scale 
* ratemarr: 5 = vry hap marr, 4 = hap than avg, 3 = avg, 2 = smewht unhap, 1 = vry unhap 
* naffairs: number of affairs within last year 
* affair: =1 if had at least one affair 
* vryhap: ratemarr == 5 
* hapavg: ratemarr == 4 
* avgmarr: ratemarr == 3 
* unhap: ratemarr == 2 
* vryrel: relig == 5 
* smerel: relig == 4 
* slghtrel: relig == 3 
* notrel: relig == 2 


### Models

Insert model exammples here...




## School Voucher Dataset

```{r, eval=T}

data( 'voucher' )
head( voucher ) %>% pander()

```






# Package 'Ecdat'

```{r, eval=F}

install.packages( "Ecdat" )
library( Ecdat )

```


## Star Dataset

* tmathssk total math scaled score 
* treadssk total reading scaled score 
* classk type of class, a factor with levels (regular,small.class,regular.with.aide) 
* totexpk years of total teaching experience 
* sex a factor with levels (boy,girl) 
* freelunk qualified for free lunch ? 
* race a factor with levels (white,black,other) 
* schidkn school indicator variable 

```{r, fig.width=10, fig.height=10 }

data( Star )
head( Star ) %>% pander()

dat <- dplyr::select( Star, tmathssk,treadssk, totexpk ) %>% 
  dplyr::rename( Math_Score=tmathssk, Reading_Score=treadssk, Teacher_Experience=totexpk )

pairs( dat, lower.panel=panel.smooth, upper.panel=panel.cor)



```



```{r}

par( mfrow=c(2,2) )
plot( Star$class, Star$tmathssk, frame.plot=F, outline=F, main="Class Type" )
plot( Star$sex, Star$tmathssk, frame.plot=F, outline=F, main="Gender" )
plot( Star$freelunk, Star$tmathssk, frame.plot=F, outline=F, main="Free Lunch" )
plot( Star$race, Star$tmathssk, frame.plot=F, outline=F, main="Race" )


```


```{r}

group.structure <- formula( Star$tmathssk ~ Star$freelunk * Star$totexpk )

boxplot( group.structure, 
         las=2, frame.plot=F, outline=F, main="Age Group by Gender",
         col=c("firebrick","steelblue"), xaxt="n",
         staplewex=0, whisklty=0, ylim=c(400,600) )

```



```{r, results='asis'}


m1 <- lm( tmathssk ~ classk, data=Star )
m2 <- lm( tmathssk ~ classk + totexpk + sex + race + freelunk, data=Star )
m3 <- lm( treadssk ~ classk, data=Star )
m4 <- lm( treadssk ~ classk + totexpk + sex + race + freelunk, data=Star )

stargazer( m1, m2, m3, m4, type="html" )


```


## Dataset 'Treatment'

```{r}

data( 'Treatment' )
head( Treatment ) %>% pander()

```

