---
title: "Difference in Difference Lab"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: no
    css: 'css/lab-instructions.css'
    includes:
      after_body: 'footer.html'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE, message=F, warning=F, fig.width = 10 )
```

```{r, echo=FALSE, results="hide"}
library(stargazer)
```


# Instructions 

This lab is loosely based on Branas et al. (2011) article, entitled ["A Difference-in-Differences Analysis of Health, Safety, and Greening Vacant Urban Space"](https://academic.oup.com/aje/article/174/11/1296/111352).


There are **4 sets of questions**. You can earn up to 100 points + bonus questions. Points are indicated next to each question. 

Remember to:

* Format your table using stargazer 
* Submit the r markdown file with all your codes. 
* Bold your answers.


## The policy problem

> **Research question**: 
> Does greening vacant lots decreases vandalism in City A? 

The 'broken windows' theory suggests that "vacant lots offer refuge to criminal and other illegal activity and visibly symbolize that a neighborhood has deteriorated, that no one is in control, and that unsafe or criminal behavior is welcome to proceed with little if any supervision" (p. 1297). 

To prevent these problems, city A has decided to implement a public program that will green some of the city's vacant lots in 4 different neighborhoods. 'Greening' includes removing trash and debris and planting grass and trees to create a park.   

Three years after its implementation, the city wants to know whether the program has been successful and propose to compare lots that have been 'greened' and lots that are still vacants within the same neighborhoods but could have been chosen for greening. 

> **Hypothesis**: 
> Greening vacant lots will decrease vandalism in City A.

## Data

Data on vandalism acts were reported by the local police and include the total number of calls about vandalism acts that the police has received during the year before and after the greening from households near the vacant lots. 

```{r, echo=TRUE}

URL <- "https://raw.githubusercontent.com/DS4PS/pe4ps-textbook/master/labs/DATA/diff-in-diff-lab.csv"
data <- read.csv( URL, stringsAsFactors=F )
 
```


 **Variable name**       | **Description    **                                   
------------------------ | ------------------------------------------------------------------------------------------------------------------------------------
Vandalism                |Number of calls about vandalism acts near to the vacant lots                      
Group                    |Treatment (=1) and control group (=0)
Post_Treatment           |Observation three years before the treatment (=0) and three years after the treatment (=1)


# Lab Questions 


## Question 1 

**Q1: Write down the difference-in-difference model** (5 points)


## Question 2

**Q2: Run the regression in R and present results in a nice table with stargazer** (5 + 5 points)

* **Q2a:** What is the average number of calls about vandalism acts in the control group before the treatment? (5 points)
* **Q2b:** Did the number of calls for vandalism in the treatment and control groups differ significantly before the treatment? Which coefficient provides you this information? (5 + 5 points)
* **Q2c:** Do treatment and control groups need to be identical in the pre-treatment period for valid inference? (10 points)
* **Q2d:** Does the number of calls in the control group change after the intervention? Specify if it increases or decreases. (5 points)
* **Q2e:** What is the average number of calls in the control group after the intervention? (5 points)
* **Q2f:** Does greening help to reduce vandalism? Report the coefficient that you need to look. (5 + 5 points)
* **Q2g:** What is the average number of calls in the treatment group after the intervention? (5 points)

## Question 3

**Q3: Let's now look at the counterfactual.** 

* **Q3a:** Write the counterfactual formula. (5 points)
* **Q3b:** Calculate the counterfactual, i.e., how many calls for vandalism acts we would expect, had the intervention not occured. (5 points)
* **Q3c:** What is the main assumption behind the counterfactual? (5 points)

Now let's imagine that you have daily data about the number of calls for vandalism that the police has received after the intervention for the following 3 years. The mayor asks another analyst to evaluate the impact of the intervention. The analyst uses an OLS grouped data regression, as illustrated in the the following model:

$$
\text{Vandalism} = b_0 + b_1 \cdot Group + b_2 \cdot Time + b_3 \cdot Group \cdot Time + e
$$
Where Group indicates the treatment (=1) and the control group (=0) and time is the time passed since the intervention (in days).


## Question 4

* **Q4a:** How do you interpret coefficient b2 and b3? (5 + 5 points)
* **Q4b:** HOw are b2 and b3 in this model different from b2 and b3 in the diff-in-diff model you estimated before? (5 + 5 points)


## BONUS QUESTIONS

* **BQ1**: Re-write the above formula imaging the starting number of calls to be different between the treatment and the control group, the overall trend over time to be positive, but no annual increase differences between the two groups exist. (5 points)

* **BQ2:** Between the diff-in-diff model and the grouped data regression, which of these is a more convincing counter-factual, in your opinion? (5 points)




---


<br>

# Submission Instructions

When you have completed your assignment, knit your RMD file to generate your rendered HTML file. 

Login to Canvas at <http://canvas.asu.edu> and navigate to the assignments tab in the course repository. Upload your HTML and RMD files to the appropriate lab submission link.


Remember to:

* name your files according to the convention: **Lab-##-LastName.Rmd**.
* show your solution, include your code.
* format your regression tables using stargazer 
* bold your answers. 
* do NOT print excessive output (like a full data set).
* follow appropriate style guidelines (spaces between arguments, etc.).


*Platforms like Canvas and Blackboard sometimes disallow you from submitting HTML files when there is embedded computer code. If this happens create a zipped folder with both the RMD and HTML files.*


<br>

-----

<br>




